---
title: 'Fig6C. Cortisol'
author: "Xuan He"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    toc_depth: 3
  html_notebook: 
    toc: true
    toc_depth: 3
---
<style type="text/css">
body{ /* Normal  */ font-size: 12px;}
td {  /* Table  */ font-size: 10px;}
h1.title {font-size: 22px; color: DarkRed; font-style: bold;}
h1 { /* Header 1 */ font-size: 16px; color: mediumblue; font-style: bold;}
h2 { /* Header 2 */ font-size: 13px; color: mediumblue; font-style: bold;}
h3 { /* Header 3 */ font-size: 11px; font-family: "Times New Roman", Times, serif; color: mediumblue;}
code.r{ /* Code block */ font-size: 12px;}
pre { /* Code block - determines code spacing between lines */ font-size: 12px;}
</style>

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/xuanhe89/Documents/WorkSpace/Research projects/Lactation/ALAC/Manuscripts/Paper #1/Figures and Tables R codes')
```


# Clean up the environment
```{r}
rm(list = ls())
```


# Load necessary libraries for data reading, plotting, and statistical analysis
```{r message=FALSE}
library(readxl)       # For reading Excel files
library(ggpubr)       # For creating publication-ready plots
library(dplyr)        # For data manipulation
library(ggh4x)        # For advanced faceting functions in ggplot2
library(rstatix)      # For statistical tests and adding significance markers
library(knitr)        # For displaying tibbles in a nice table format
library(kableExtra)   # For enhancing table formatting
```


# Load data
```{r}
# Load hormonal data
PP_hormones <- read_excel("Data/Data.xlsx", sheet = "PP_hormones")
Nec_hormones <- read_excel("Data/Data.xlsx", sheet = "Nec_hormones")

# Load pig information
PigInfo <- read_excel("Data/Data.xlsx", sheet = "PigInfo")
```

# Data organization
```{r}
PP.data <- PP_hormones %>% inner_join(PigInfo, by = "PigID") %>%
                             mutate(Day = "Day 11 or 12") %>%
                             select(Diet, Day, Cortisol, TimePoint)

Nec.data <- Nec_hormones %>% inner_join(PigInfo, by = "PigID") %>%
              mutate(Day = "Day 16") %>%
              mutate(TimePoint = "") %>%
              select(Diet, Day, Cortisol, TimePoint)

```

```{r}
# Combine postprandial and necropsy data
All.data <- bind_rows(PP.data, Nec.data) %>% 
              mutate(Diet = factor(Diet, levels = c("ALAC", "WPI", "SF"))) %>% 
              mutate(TimePoint = factor(TimePoint, levels = c("baseline", "30 min", "60 min", "120 min")))
```

# Plotting cortisol levels by diet and time point
```{r fig.width=6, fig.height=4}
Fig <- All.data %>% ggerrorplot(x = "TimePoint", y = "Cortisol", color = "Diet", 
                      palette = c("turquoise3","purple", "orange"),
                      na.rm = TRUE, xlab = "", 
                      ylab ="Cortisol (ug/L)",  add = "mean_se", position = position_dodge(0.6)) +
                    facet_nested(.~ Day , scales = "free", space = "free_x") + 
                    theme_classic() 

Fig

# Save the figure in pdf format
ggsave(plot=Fig, height=4, width=6, dpi=300, filename="Figure 6/Fig6C.pdf", useDingbats=FALSE)
```

# Group difference evaluation
```{r}
All.data %>% group_by(TimePoint) %>%
               wilcox_test(Cortisol ~ Diet, p.adjust.method = "none", paired = FALSE) %>% # Perform Wilcoxon test (Mann-Whitney U test)
               select(-p.adj) %>%
               add_significance(cutpoints = c(0, 0.05, 0.1,  1), symbols = c("*", "#", "ns")) %>%
               select(TimePoint, group1, group2, p, p.signif) %>%
               kable() %>% # Print the tibble in a nicely formatted table
               kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting with kableExtra
```


```{r}
sessionInfo()
```