---
title: 'Fig5C. Serotonin'
author: "Xuan He"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    toc_depth: 3
  html_notebook: 
    toc: true
    toc_depth: 3
---
<style type="text/css">
body{ /* Normal  */ font-size: 12px;}
td {  /* Table  */ font-size: 10px;}
h1.title {font-size: 22px; color: DarkRed; font-style: bold;}
h1 { /* Header 1 */ font-size: 16px; color: mediumblue; font-style: bold;}
h2 { /* Header 2 */ font-size: 13px; color: mediumblue; font-style: bold;}
h3 { /* Header 3 */ font-size: 11px; font-family: "Times New Roman", Times, serif; color: mediumblue;}
code.r{ /* Code block */ font-size: 12px;}
pre { /* Code block - determines code spacing between lines */ font-size: 12px;}
</style>

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/xuanhe89/Documents/WorkSpace/Research projects/Lactation/ALAC/Manuscripts/Paper #1/Figures and Tables R codes')
```


# Clean up the environment
```{r}
rm(list = ls())
```


# Load library
```{r message=FALSE}
library(readxl)       # For reading Excel files
library(ggpubr)       # For creating publication-ready plots
library(dplyr)        # For data manipulation
library(tidyr)        # For reshaping data
library(ggh4x)        # For advanced faceting functions in ggplot2
library(rstatix)      # For statistical tests and adding significance markers
library(knitr)        # For displaying tibbles in a nice table format
library(kableExtra)   # For enhancing table formatting
```


# Load data
```{r}
# Load targeted serum, liver and brain data
Serum <- read_excel("Data/Data.xlsx", sheet = "Nec_serum_targeted")
Liver <- read_excel("Data/Data.xlsx", sheet = "Liver_targeted")
Brain <- read_excel("Data/Data.xlsx", sheet = "Brain_targeted")

# Load pig information and necropsy metadata
PigInfo <- read_excel("Data/Data.xlsx", sheet = "PigInfo")
Nec_metadata <- read_excel("Data/Data.xlsx", sheet = "Nec_metadata")
```


# Data organization

```{r}
Data.serum <- inner_join(Serum, PigInfo, by = "PigID") %>%
               select(Diet, Serotonin) %>%
               mutate(Region = "Serum") %>%
               pivot_longer(cols = Serotonin, names_to = "variable", values_to = "value") %>%
               mutate(Class = "I")


Data.liver <- inner_join(Liver, PigInfo, by = "PigID") %>%
               select(Diet, Serotonin) %>%
               mutate(Region = "Liver") %>%
               pivot_longer(cols = Serotonin, names_to = "variable", values_to = "value") %>%
               mutate(Class = "II")

Data.brain <- inner_join(Brain, PigInfo, by = "PigID") %>%
               select(Diet, Region, Serotonin) %>%
               pivot_longer(cols = Serotonin, names_to = "variable", values_to = "value") %>%
               mutate(Class = "II")
```

```{r}
# Combine serum, liver, and brain data
All.data <- bind_rows(Data.serum, Data.liver, Data.brain) %>%
              mutate(Region = factor(Region, levels = c("Serum", "Liver", "Hippocampus", "Hypothalamus", "PFC", "Striatum")))
```

# Plotting data for serotonin concentration by diet and sample type
```{r fig.width=7.5, fig.height=4}
Fig <- All.data %>% 
          mutate(Diet = factor(Diet, levels = c("ALAC", "WPI", "SF"))) %>%
          ggerrorplot(x = "Region", y = "value", color = "Diet", 
                      palette = c("turquoise3","purple", "orange"),
                      na.rm = TRUE, xlab = "", 
                      ylab ="Serotonin concentration",  add = "mean_se", position = position_dodge(0.7)) +
          facet_grid2(.~ Class , scales = "free", independent = "y", space = "free_x") + 
          theme_classic() +
          theme(strip.background = element_blank(), strip.text.x = element_blank())


Fig

# Save the figure in pdf format:
ggsave(plot=Fig, height=4, width=7.5, dpi=300, filename="Figure 5/Fig5C.pdf", useDingbats=FALSE)
```

# Group difference evaluation
```{r}
#This function transforms the input values by the generalized #log function.
glog <- function(y) {
  #Using lambda = 1 
  yt <- log(y+sqrt(y^2+1))
  return(yt)
}
```

## ALAC vs WPI: Serum and liver data, adjusting for intake volume and postprandial time
```{r}
Serum %>% inner_join(PigInfo, by = "PigID") %>%
          inner_join(Nec_metadata, by = "PigID") %>%
            filter(Diet %in% c("ALAC", "WPI")) %>%
            mutate(glog.5HT = glog(Serotonin * 10)) %>% # glog transformation
            anova_test(glog.5HT ~ Diet + SowID + Formula.consumed.g + Time.postprandial.min, type = 1)
```
```{r}
Liver %>% inner_join(PigInfo, by = "PigID") %>%
          inner_join(Nec_metadata, by = "PigID") %>%
            filter(Diet %in% c("ALAC", "WPI")) %>%
            mutate(glog.5HT = glog(Serotonin * 10)) %>% # glog transformation
            anova_test(glog.5HT ~ Diet + SowID + Formula.consumed.g + Time.postprandial.min, type = 1)
```


## Between formula-fed groups and SF, adjusting for sow ID

```{r}
Serum %>% inner_join(PigInfo, by = "PigID") %>%
            mutate(glog.5HT = glog(Serotonin * 10)) %>% # glog transformation
            tukey_hsd(glog.5HT ~ Diet + SowID) %>% # Perform Tukey HSD test
            filter(term == "Diet") %>%
            add_significance(p.col = "p.adj", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>%
            kable() %>% # Print the tibble in a nicely formatted table
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting with kableExtra
```


```{r}
Liver %>% inner_join(PigInfo, by = "PigID") %>%
            mutate(glog.5HT = glog(Serotonin * 10)) %>% # glog transformation
            tukey_hsd(glog.5HT ~ Diet + SowID) %>% # Perform Tukey HSD test
            filter(term == "Diet") %>%
            add_significance(p.col = "p.adj", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>%
            kable() %>% # Print the tibble in a nicely formatted table
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting with kableExtra
```


## Brain data, adjusting for sow ID

```{r}
Brain %>% inner_join(PigInfo, by = "PigID") %>%
            mutate(glog.5HT = glog(Serotonin * 10)) %>% # Applying glog transformation
            group_by(Region) %>%
            tukey_hsd(glog.5HT ~ Diet + SowID) %>% # Perform Tukey HSD test
            filter(term == "Diet") %>%
            add_significance(p.col = "p.adj", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>%
            filter(p.adj.signif != "ns") %>%
            select(Region, group1, group2, p.adj, p.adj.signif) %>%
            kable() %>% # Print the tibble in a nicely formatted table
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting with kableExtra
```


```{r}
sessionInfo()
```