---
title: 'Fig5B. Ketone bodies'
author: "Xuan He"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    toc_depth: 3
  html_notebook: 
    toc: true
    toc_depth: 3
---
<style type="text/css">
body{ /* Normal  */ font-size: 12px;}
td {  /* Table  */ font-size: 10px;}
h1.title {font-size: 22px; color: DarkRed; font-style: bold;}
h1 { /* Header 1 */ font-size: 16px; color: mediumblue; font-style: bold;}
h2 { /* Header 2 */ font-size: 13px; color: mediumblue; font-style: bold;}
h3 { /* Header 3 */ font-size: 11px; font-family: "Times New Roman", Times, serif; color: mediumblue;}
code.r{ /* Code block */ font-size: 12px;}
pre { /* Code block - determines code spacing between lines */ font-size: 12px;}
</style>

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/xuanhe89/Documents/WorkSpace/Research projects/Lactation/ALAC/Manuscripts/Paper #1/Figures and Tables R codes')

```


# Clean up the environment
```{r}
rm(list = ls())
```


# Load library
```{r message=FALSE, warning=FALSE}
library(readxl)       # For reading Excel files
library(ggpubr)       # For creating publication-ready plots
library(dplyr)        # For data manipulation
library(tidyr)        # For reshaping data, replacing melt() from reshape2
library(ggh4x)        # For advanced faceting functions in ggplot2
library(rstatix)      # For statistical tests and adding significance markers
library(knitr)        # For displaying tibbles in a nice table format
library(kableExtra)   # For enhancing table formatting
```


# Load data
```{r}
# Load targeted serum and liver data
Serum <- read_excel("Data/Data.xlsx", sheet = "Nec_serum_targeted")
Liver <- read_excel("Data/Data.xlsx", sheet = "Liver_targeted")

# Load pig information and necropsy metadata
PigInfo <- read_excel("Data/Data.xlsx", sheet = "PigInfo")
Nec_metadata <- read_excel("Data/Data.xlsx", sheet = "Nec_metadata")
```


# Data organization
```{r}
# Define the list of metabolites for each sample type
ls.serum.metabolites <- c("3-Hydroxybutyrate", "Acetoacetate")
ls.liver.metabolites <- c("3-Hydroxybutyrate")
```

```{r}
Data.serum <- inner_join(Serum, PigInfo, by = "PigID") %>%
                select(Diet, all_of(ls.serum.metabolites)) %>%
                mutate(Region = "Serum") %>%
                pivot_longer(cols = all_of(ls.serum.metabolites), names_to = "variable", values_to = "value")


Data.liver <- inner_join(Liver, PigInfo, by = "PigID") %>%
                select(Diet, all_of(ls.liver.metabolites)) %>%
                mutate(Region = "Liver") %>%
                pivot_longer(cols = all_of(ls.liver.metabolites), names_to = "variable", values_to = "value")
```

```{r}
# Combine serum and liver data into a single dataset
All.data <- bind_rows(Data.serum, Data.liver) %>%
              mutate(Region = factor(Region, levels = c("Serum", "Liver")))
```

# Plotting data
```{r fig.width=6, fig.height=4}
Fig <- All.data %>% 
        mutate(Diet = factor(Diet, levels = c("ALAC", "WPI", "SF"))) %>%
        ggerrorplot(x = "variable", y = "value", color = "Diet", 
            palette = c("turquoise3", "purple", "orange"),
            na.rm = TRUE, 
            xlab = "", 
            ylab = "Concentration", 
            add = "mean_se", 
            size = 0.6,
            position = position_dodge(0.7)) +
            facet_nested(~Region + variable, scales = "free", independent = "y") +  # Free x axis with proportional facet widths using facet_nested
            theme_classic2() +
            theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())


Fig

# Save the figure in pdf format:
ggsave(plot=Fig, height=4, width=6, dpi=300, filename="Figure 5/Fig5B.pdf", useDingbats=FALSE)
```


# Group difference evaluation
```{r}
#This function transforms the input values by the generalized #log function.
glog <- function(y) {
  #Using lambda = 1 
  yt <- log(y+sqrt(y^2+1))
  return(yt)
}
```

##ALAC vs WPI: Serum and liver data, adjusting for intake volume and postprandial time

```{r}
Serum %>% inner_join(PigInfo, by = "PigID") %>%
            inner_join(Nec_metadata, by = "PigID") %>%
            select(`3-Hydroxybutyrate`, `Acetoacetate`, Diet, SowID, Formula.consumed.g, Time.postprandial.min) %>%
            pivot_longer(cols = c(`3-Hydroxybutyrate`, `Acetoacetate`), names_to = "variable", values_to = "value") %>%
            filter(Diet %in% c("ALAC", "WPI")) %>%
            group_by(variable) %>%
            mutate(glog.value = glog(value * 10)) %>% # Apply generalized log transformation to metabolite values
            anova_test(glog.value ~ Diet + SowID + Formula.consumed.g + Time.postprandial.min, type = 1) %>% # Perform ANOVA
            add_significance(p.col = "p", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>% # Add significance markers
            as_tibble() %>%
            filter(Effect == "Diet") %>%
            kable() %>% # Print the tibble in a nicely formatted table
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting with kableExtra
```

```{r}
Liver %>% inner_join(PigInfo, by = "PigID") %>%
            inner_join(Nec_metadata, by = "PigID") %>% 
            filter(Diet %in% c("ALAC", "WPI")) %>%
            mutate(glog.3HB = glog(`3-Hydroxybutyrate` * 10)) %>% # glog transformation
            anova_test(glog.3HB ~ Diet + SowID + Formula.consumed.g + Time.postprandial.min, type = 1) 
```

# Between formula-fed groups and SF, adjusting for sow ID

```{r}
Serum %>% inner_join(PigInfo, by = "PigID") %>%
            inner_join(Nec_metadata, by = "PigID") %>%
            select(`3-Hydroxybutyrate`, `Acetoacetate`, Diet, SowID) %>%
            pivot_longer(cols = c(`3-Hydroxybutyrate`, `Acetoacetate`), names_to = "variable", values_to = "value") %>%
            group_by(variable) %>%
            mutate(glog.value = glog(value * 10)) %>% # Apply generalized log transformation to metabolite values
            tukey_hsd(glog.value ~ Diet + SowID) %>% # Perform Tukey HSD test
            filter(term == "Diet") %>%
            add_significance(p.col = "p.adj", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>%
            kable() %>% # Print the tibble in a nice table format
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting
```


```{r}
Liver %>% inner_join(PigInfo, by = "PigID") %>%
            inner_join(Nec_metadata, by = "PigID") %>%
            select(`3-Hydroxybutyrate`, Diet, SowID) %>%
            pivot_longer(cols = c(`3-Hydroxybutyrate`), names_to = "variable", values_to = "value") %>%
            group_by(variable) %>%
            mutate(glog.value = glog(value * 10)) %>%
            tukey_hsd(glog.value ~ Diet + SowID) %>%
            filter(term == "Diet") %>%
            add_significance(p.col = "p.adj", cutpoints = c(0, 0.05, 0.1, 1), symbols = c("*", "#", "ns")) %>%
            kable() %>% # Print the tibble in a nice table format
            kable_styling(bootstrap_options = c("striped", "hover")) # Enhance table formatting

```

```{r}
sessionInfo()
```